---
title: "Covid19 Vaccine Study Final2023"
author: "Tyler Hamblin"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")

```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_0$ : The data for the preventative vaccine has no statistical significance to changing infection rate across the listed sub populations. 


$H_A$ : The data for the preventative vaccine does have statistical significance in how it prevents infection for each sub group. 



# Methods
According to the data set this is going to be a categorical situation explained by another categorical variable. After creating separate data sets for each of the sub populations we will use the CandP helper files to graphically display bar charts, pull xtabs commands for numerical data, run Chi-Squared for inferential data, and end with Fisher code to declare a confidence interval. Since we are limiting this investigation to two types of data there are no loops required to measure multiple variables. 

# Results

## Males

### Graphical Results
```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```


The above barchart compares the graphical results for only the male sub group from the data set. It displays the rates of contracting Covid-19 explained by whether or not the individual was vaccinated. The left side of the chart accounts for those with positive test results and the right shows people who did not contract it. More than twice the amount of people in the Covid-19 category only had the placebo, suggesting that the vaccine does have efficacy for this group. 

### Numerical Results
```{r}
tableMale <- xtabs(~infected + treatment, data=male)
rowPerc(tableMale)
colPerc(tableMale)
```


These numerical results confirm the suggestions from the barchart that people who have the vaccine tend to contract Covid-19 half as much as the people who only received the placebo. The placebo got it about 70% of the time and the vaccinated group got it 30%. Again our assumption is that the vaccination is effective for males. 

### Inferential Results
```{r}
chisq.test(tableMale)
chisqtestGC(tableMale)
maleFisher <- fisher.test(tableMale)
maleFisher
```



The Chi Squared test produces a p-value that is significantly smaller than the suggested 0.05 limit. Coming in at $1.098\times10^{-13}$ , there is even greater evidence that a statistical significance is responsible for the data. The Chi-squared statistic is 55 which comes in larger than the degrees of freedom for this data. The chance of an unvaccinated male getting Covid-19 are 2.2 times higher than a male with the vaccine. 

### Confidence Interval
```{r}
oddsMale = maleFisher$estimate/(maleFisher$estimate+1)*100
print(oddsMale)
maleLow=(maleFisher$conf.int[1]/(maleFisher$conf.int[1]+1))*100
maleHigh=(maleFisher$conf.int[2]/(maleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Males: ",maleLow," to ",maleHigh)
```


The confidence interval for this data set is >50% which meets the typical vaccine threshold of >60%. This is enough data to confidently say that the vaccine is a good option for preventing the contraction of Covid-19 for males. 

## Females

### Graphical Results
```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")
```


This barchart solely depicts the graphical results for all the females within this data set. Again the comparison shows the contraction of Covid-19 based on vaccination status. The left is for those that have contracted the virus and the right is for those that have not. The placebo has about one third more Covid-19 positive females compared to the vaccinated women on that side. 

### Numerical Results
```{r}
tableFemale <- xtabs(~infected + treatment, data=female)
rowPerc(tableFemale)
colPerc(tableFemale)
```


These numerical results again confirm the suggestion of the graphical data that more people get Covid-19 who have only recieved the placebo treatment. Approximately 60% of the time the unvaccinated people were infected where as only 40% of the time the vaccinated people tested positive. The vaccine is still generally affective for the female sub group but slightly less than the male category. 

### Inferential Results
```{r}
chisq.test(tableFemale)
chisqtestGC(tableFemale)
femaleFisher <- fisher.test(tableFemale)
femaleFisher
```


From the Chi-Squared test, we get a p-value of $2.732\times10^{-4}$ that is below the 0.05 limit showing there is statistical significance. The Chi-Squared statistic is 13, also higher than the degrees of freedom for this instance. An unvaccinated female will get Covid-19 1.5 times more than a woman who did receive the vaccine.

### Confidence Interval
```{r}
oddsFemale = femaleFisher$estimate/(femaleFisher$estimate +1)*100
print(oddsFemale)
femaleLow=(femaleFisher$conf.int[1]/(femaleFisher$conf.int[1]+1))*100
femaleHigh=(femaleFisher$conf.int[2]/(femaleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Females: ",femaleLow," to ",femaleHigh)
```


Once again there is a >50% confidence interval, lower than the >60% suggestion, so the vaccination is still advisable for preventing infection for females. 

## LGBTQ

### Graphical Results
```{r}
barchartGC(~infected + treatment,data=lgbtq)
barchartGC(~infected + treatment,data=lgbtq, type="percent")
```

Above the barchart compares the relationship of the placebo and vaccine categories for people only from the LGBTQ category. In response to how many more people get Covid-19 it appears that the vaccine makes you about 5 times more likely as prone to it. The left side of the graph shows those that do have Covid-19 with the right side showing the data for those testing negative. 


### Numerical Results
```{r}
tableLGBTQ <- xtabs(~infected + treatment, data=lgbtq)
rowPerc(tableLGBTQ)
colPerc(tableLGBTQ)
```

The numerical results are a reflection of the graphical data that show only about 15% of the Covid-19 positive people had the placebo treatment and instead about 85% of them were vaccinated. This is the first instance where the vaccine is not effective and most likely worsens a patients chance at avoiding the disease. For an unknown reason being apart of the LGBTQ group reverses the success rates of this treatment method. 

### Inferential Results
```{r}
chisq.test(tableLGBTQ)
chisqtestGC(tableLGBTQ)
lgbtqFisher <- fisher.test(tableLGBTQ)
lgbtqFisher
```


The p-value of $6.668\times10^{-8}$ comes in well under the suggested 0.05 threshold, showing that this reverse effect of the vaccine is not a fluke. The Chi-Squared statistic, 29, supports this as it is larger than the degrees of freedom. The odds for an unvaccinated LGBTQ person getting Covid-19 compared to a vaccinated one are about 1:5. 

### Confidence Interval
```{r}
oddsLGBTQ = lgbtqFisher$estimate/(lgbtqFisher$estimate +1)*100
print(oddsLGBTQ)
lgbtqLow=(lgbtqFisher$conf.int[1]/(lgbtqFisher$conf.int[1]+1))*100
lgbtqHigh=(lgbtqFisher$conf.int[2]/(lgbtqFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for LGBTQ: ",lgbtqLow," to ",lgbtqHigh)
```

The Confidence interval for LGBTQ members is only about 8%-27% which is far lower than the >50% minimum so this vaccine does not serve the purpose of limiting infection rate for Covid-19 for this sub group. 

## Druggies

### Graphical Results
```{r}
barchartGC(~infected + treatment,data=drugUser)
barchartGC(~infected + treatment,data=drugUser, type="percent")
```

The barchart above now shows the graphical results for all people identified on drugs and how they contract Covid-19 based on whether or not they have the vaccine. Left shows positive right shows negative. In this instance it was five or six times more likely that a person who had Covid-19 had also received the vaccine rather than the placebo. This could suggest that drug users are already much more pre disposed to contracting Covid-19 because of weakened immune systems or because of a reverse affect the vaccination has with the drugs in a persons system. This would require a separate study. 


### Numerical Results
```{r}
tableDrug <- xtabs(~infected + treatment, data=drugUser)
rowPerc(tableDrug)
colPerc(tableDrug)
```

According to the numerical results the placebo group accounted for 12% of the people who contracted Covid-19 and the vaccinated people accounted for the other 88%. This confirms the notion from the barchart that with drug users, as opposed to males and females, the vaccine actually increases how often they get Covid-19. The vaccine so far is not effective or advisable. 


### Inferential Results
```{r}
chisq.test(tableDrug)
chisqtestGC(tableDrug)
drugFisher <- fisher.test(tableDrug)
drugFisher
```


The p-value of $2.842\times10^{-9}$ still suggests a statistical difference within the data so the reverse in effectiveness is not mere chance. The Chi-Squared statistic is 35 still larger than the degrees of freedom. The odds of a placebo treated versus vaccinated drug user contracting the virus is about 1:8. 



### Confidence Interval
```{r}
oddsDrug = drugFisher$estimate/(drugFisher$estimate +1)*100
print(oddsDrug)
drugLow=(drugFisher$conf.int[1]/(drugFisher$conf.int[1]+1))*100
drugHigh=(drugFisher$conf.int[2]/(drugFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Drug Users: ",drugLow," to ",drugHigh)
```


The confidence interval for drug users is only 5-22% which is much lower than the >50% recommendation so this should not be implemented as a reliable precaution for the spread of Covid-19 amongst drug users. 


# Overall Results and Conclusions






